Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-12 doc/manuscript/ESSDmanuscript.Rmd
draft conclusions

**Abstract.**
Forests are critical for climate change mitigation and constitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
We begin by reconciling terminology and mapping ForC fields into EFDB.
This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
As of `r format(Sys.time(), '%B %d, %Y')`, ForC contained ~`r n_ind_records_EFDB_variables` independent records that would be relevant to EFDB, `r n_records_sent_to_EFDB` of which have been submitted to date.
Among the data in ForC, there is disproportionate representation of biomass (particularly aboveground) stocks, with far fewer records for dead organic matter and soil C, and relatively few or no records for net annual increments or C fluxes into (gains) or out of (losses) the IPCC-defined C pools.
Geographic representation is also quite uneven, with the highest densities of relevant records in temperate forests, and with relatively scant representation of tropical forests in Africa and Asia. 
ForC represents a diversity of stand ages, although records for young stands are primarily limited to C stocks, as opposed to net increments of fluxes.
This distribution of records is generally reflected in the subset of records that have been submitted to EFDB to date.
In the future, forest C estimates in EFDB can be improved through targeted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.
Given that climate change is rapidly impacting the world's forests, timely reporting of recent estimates will be especially critical to accurate forest C accounting.

@@

Once EFDB-relevant data are available in peer-reviewed publications, they may be submitted directly to EFDB or may use the ForC - EFDB data pipeline developed here.
For individual publications, the former option will generally be more efficient.
However, data incorporated into ForC as well as EFDB will be more broadly useful; for example, these data may be used for basic science [e.g., @banburymorgan_global_2021; @anderson-teixeira_carbon_2021], analyses of forest-based climate change mititagation potential [e.g., @cook-patton_mapping_2020; @goldstein_protecting_2020], and model benchmarking [@fer_ecosystem_2021].

# Conclusions

<!--The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.-->

An important challenge is that forests are changing rapidly, and data collected a decade ago may no longer be relevant, particularly in the cases of C increments and fluxes.

>>>>

**Abstract.**
Forests are critical for climate change mitigation and constitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
We begin by reconciling terminology and mapping ForC fields into EFDB.
This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
As of `r format(Sys.time(), '%B %d, %Y')`, ForC contained ~`r n_ind_records_EFDB_variables` independent records relevant to EFDB, `r n_records_sent_to_EFDB` of which have been submitted to date.
Among the data in ForC, there is disproportionate representation of biomass (particularly aboveground) stocks, with far fewer records for dead organic matter and soil C, and relatively few or no records for net annual increments or C fluxes into (gains) or out of (losses) the IPCC-defined C pools.
Geographic representation is also quite uneven, with the highest densities of relevant records in temperate forests, and with relatively scant representation of tropical forests in Africa and Asia. 
This distribution of records is generally reflected in the subset of records that have been submitted to EFDB to date.
In the future, forest C estimates in EFDB can be improved through targeted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.
Given that climate change is rapidly impacting the world's forests, timely reporting of recent estimates will be especially critical to accurate forest C accounting.

@@

Once EFDB-relevant data are available in peer-reviewed publications, they may be submitted directly to EFDB or may use the ForC - EFDB data pipeline developed here.
For individual publications, the former option will generally be more efficient.
However, data incorporated into ForC as well as EFDB will be more broadly useful; for example, these data may be used for basic science [e.g., @banburymorgan_global_2021; @anderson-teixeira_carbon_2021], analyses of forest-based climate change mitigation potential [e.g., @cook-patton_mapping_2020; @goldstein_protecting_2020], and model benchmarking [@fer_ecosystem_2021].

# Conclusions
<!--The conclusion goes here.
You can modify the section name with  `\conclusions[modified heading if necessary]`.-->

The ForC database contains large numbers of records that could potentially be useful for C accounting under IPCC guidelines.
Here we have developed a framework for transferring these records to the EFDB, making them more accessible for forest C accounting under IPCC guidelines [@ipcc_2006_2006;@ipcc_2019_2019].
As of `r format(Sys.time(), '%B %d, %Y')`, we have submitted `r n_records_sent_to_EFDB` records to EFDB.
Although this represents just `r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables*100,0)`% of relevant records in ForC, it substantially increases the number of forest land records in EFDB.
The records transferred to EFDB and present in ForC are very unevenly distributed across variables, regions, and forest types (Figs. `r fig_map`-`r fig_histograms`, Table `r table_variables`), reflecting broader patterns in allocation of research effort.

Going forward, forest researchers can make their research more useful for forest C accounting under IPCC guidelines by calculating and reporting results in ways that are compatible with IPCC guidelines (Tables `r table_pools`, `r table_recommendations`).
In addition, substantial investments in research and researchers focused on ground-based measurement of forest C will be required to fill knowledge gaps and thereby increase the accuracy of forest C accounting. 
This challenge is heightened by the fact that forests are changing rapidly [@refs], and data collected a decade or more in the past may no longer be relevant.
This heightens the need for an efficient system of making forest C data accessible for national greenhouse gas inventories.
We view this publication as one step towards that goal.

### 2023-05-12 doc/manuscript/ESSDmanuscript.tex
draft conclusions

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). As of May 12, 2023, ForC contained
\textasciitilde19323 independent records that would be relevant to EFDB,
1438 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. ForC represents a
diversity of stand ages, although records for young stands are primarily
limited to C stocks, as opposed to net increments of fluxes. This
distribution of records is generally reflected in the subset of records
that have been submitted to EFDB to date. In the future, forest C
estimates in EFDB can be improved through targeted research to fill
critical gaps, reporting of information required by IPCC, and continued
submission of data from scientific publications to the EFDB. Given that
climate change is rapidly impacting the world's forests, timely
reporting of recent estimates will be especially critical to accurate
forest C accounting.

@@

Once EFDB-relevant data are available in peer-reviewed publications,
they may be submitted directly to EFDB or may use the ForC - EFDB data
pipeline developed here. For individual publications, the former option
will generally be more efficient. However, data incorporated into ForC
as well as EFDB will be more broadly useful; for example, these data may
be used for basic science
\citep[e.g.,][]{banburymorgan_global_2021, anderson-teixeira_carbon_2021},
analyses of forest-based climate change mititagation potential
\citep[e.g.,][]{cook-patton_mapping_2020, goldstein_protecting_2020},
and model benchmarking \citep{fer_ecosystem_2021}.

@@

An important challenge is that forests are changing rapidly, and data
collected a decade ago may no longer be relevant, particularly in the
cases of C increments and fluxes.

>>>>

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). As of May 12, 2023, ForC contained
\textasciitilde19323 independent records relevant to EFDB, 1438 of which
have been submitted to date. Among the data in ForC, there is
disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C accounting.

@@

Once EFDB-relevant data are available in peer-reviewed publications,
they may be submitted directly to EFDB or may use the ForC - EFDB data
pipeline developed here. For individual publications, the former option
will generally be more efficient. However, data incorporated into ForC
as well as EFDB will be more broadly useful; for example, these data may
be used for basic science
\citep[e.g.,][]{banburymorgan_global_2021, anderson-teixeira_carbon_2021},
analyses of forest-based climate change mitigation potential
\citep[e.g.,][]{cook-patton_mapping_2020, goldstein_protecting_2020},
and model benchmarking \citep{fer_ecosystem_2021}.

@@

The ForC database contains large numbers of records that could
potentially be useful for C accounting under IPCC guidelines. Here we
have developed a framework for transferring these records to the EFDB,
making them more accessible for forest C accounting under IPCC
guidelines \citep{ipcc_2006_2006, ipcc_2019_2019}. As of May 12, 2023,
we have submitted 1438 records to EFDB. Although this represents just
7\% of relevant records in ForC, it substantially increases the number
of forest land records in EFDB. The records transferred to EFDB and
present in ForC are very unevenly distributed across variables, regions,
and forest types (Figs. 2-3, Table 2), reflecting broader patterns in
allocation of research effort.

Going forward, forest researchers can make their research more useful
for forest C accounting under IPCC guidelines by calculating and
reporting results in ways that are compatible with IPCC guidelines
(Tables 1, 3). In addition, substantial investments in research and
researchers focused on ground-based measurement of forest C will be
required to fill knowledge gaps and thereby increase the accuracy of
forest C accounting. This challenge is heightened by the fact that
forests are changing rapidly \citep{refs}, and data collected a decade
or more in the past may no longer be relevant. This heightens the need
for an efficient system of making forest C data accessible for national
greenhouse gas inventories. We view this publication as one step towards
that goal.
